---
title: "Chapter2"
output: html_document
date: "2024-10-02"
---
Install packages etc.
```{r}
library(rethinking)
library(ggplot2)
library(cowplot)
```


R code example 2.3 - 2.4
```{r}
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )
# define prior
prior <- rep( 1 , 20 )
# compute likelihood at each value in grid
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
# 2.4
plot( p_grid , posterior , type="b" ,
xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )
```
2.5
cutoff pgrid at 0.5
```{r}
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior , type="b" ,
xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )

```
more messing with the prior
```{r}
prior <- exp( -5*abs( p_grid - 0.5 ) )
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior , type="b" ,
xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )
```

### Exercises

2M1. Recall the globe tossing model from the chapter. Compute and plot the grid approximate
posterior distribution for each of the following sets of observations. In each case, assume a uniform prior for p.
(1) W, W, W
(2) W, W, W, L
(3) L, W, W, L, W, W, W

```{r}
p_grid <- seq( from=0 , to=1 , length.out=20 )
# define prior
prior <- rep( 1 , 20 )# compute likelihood at each value in grid
likelihood <- dbinom( 3, size=3 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot( p_grid , posterior , type="b" ,
xlab="probability of water 1" , ylab="posterior probability" )
mtext( "20 points" )

df1<-data.frame(p_grid, posterior)
a<-ggplot(df1,aes(x=p_grid, y=posterior)) + geom_point() +
  geom_line()
#2
likelihood <- dbinom( 3, size=4 , prob=p_grid )
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
df2<-data.frame(p_grid, posterior)
b<-ggplot(df2,aes(x=p_grid, y=posterior)) + geom_point() +
  geom_line()
#3
likelihood <- dbinom( 5, size=7 , prob=p_grid )
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
df3<-data.frame(p_grid, posterior)
c<-ggplot(df3,aes(x=p_grid, y=posterior)) + geom_point() +
  geom_line()

plot_grid(a,b,c, labels = c("1","2","3"),ncol = 2)
```
2M2.
Now assume a prior for p that is equal to zero when p < 0.5 and is a positive constant when
p ≥0.5. Again compute and plot the grid approximate posterior distribution for each of the sets of
observations in the problem just above.
```{r}
p_grid <- seq( from=0 , to=1 , length.out=20 )
prior <- ifelse( p_grid < 0.5 , 0 , 1 )

```

2M3. 
Suppose there are two globes, one for Earth and one for Mars. The Earth globe is 70% covered
in water. The Mars globe is 100% land. Further suppose that one of these globes—you don’t know
which—was tossed in the air and produced a “land” observation. Assume that each globe was equally
likely to be tossed. Show that the posterior probability that the globe was the Earth, conditional on seeing “land” (Pr(Earth|land)), is 0.23.
```{r}

```

2M4. Suppose you have a deck with only three cards. Each card has two sides, and each side is either
black or white. One card has two black sides. The second card has one black and one white side. The
third card has two white sides. Now suppose all three cards are placed in a bag and shuffled. Someone
reaches into the bag and pulls out a card and places it flat on a table. A black side is shown facing up,
but you don’t know the color of the side facing down. Show that the probability that the other side is
also black is 2/3. Use the counting method (Section 2 of the chapter) to approach this problem. This
means counting up the ways that each card could produce the observed data (a black side facing up
on the table).

```{r}

```
2H1. Suppose there are two species of panda bear. Both are equally common in the wild and live
in the same places. They look exactly alike and eat the same food, and there is yet no genetic assay
capable of telling them apart. They differ however in their family sizes. Species A gives birth to twins
10% of the time, otherwise birthing a single infant. Species B births twins 20% of the time, otherwise
birthing singleton infants. Assume these numbers are known with certainty, from many years of field
research.
Now suppose you are managing a captive panda breeding program. You have a new female panda
of unknown species, and she has just given birth to twins. What is the probability that her next birth
will also be twins?

pr(A|B) = Pr(B|A)pr(A)/Pr(B)
-------
A, 0.5 - 10% twins
B, 0.5 - 20% twins
probability that she is species A

pr(A|twins) = pr(twins|A)Pr(A)/Pr(twins)

probability of twins
pr(twins) = 0.1x0.5 + 0.2x0.5
Pr(A|twins) = 0.10x0.5/0.15 = 0.33
Pr(B|twins) = 1-0.33 = 0.66
______

Pr twins the second time, add in new probability
Pr(twins2) = 0.1x0.33 + 0.2x0.66 = 0.167

_____________
singletons second time
A -> panda is spp A
B -> panda is spp B
Ti -> having twins in birth i

Pr(A|t1 and not t2) = P(not twins2 and twins1|A)xP(A)/Pr(not twins2 and twins1)
P(not twins2 and twins1|A) = Pr(T1|A)xPr(notTwins2|A) = 1/10x9/10 = 9/100
Pr(A) = 0.5
Pr(B)xPr(not twins2 and twins1|B)=0.2 x 0.8 = 16/100
Pr(not twins2 AND twins1) = Pr(A)xPr(not twins2 and twins1|A) +Pr(B)xPr(not twins2 and twins1|B)
= 1/2x9/100 / 1/8 = 0.36



